---
title: "COVID-19 Mitigation vs. Suppression Social Distancing"
author:
- name: Sarah Nowak, Pedro de Lima, Raffaele Vardavas 
  affiliation: RAND Corporation
#bibliography: ../../WriteUp/COVID-19.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: html_notebook
output:
  rmdformats::readthedown
pandoc_args: --natbib
biblio-style: plain
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
# Needed packages to perform the analysis
load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
} 

packages <- c("knitr","kableExtra","rmdformats","rmarkdown","dplyr","here","DiagrammeR",
              "ggplot2","scales","ggthemes","bvpSolve","deSolve","class","pracma")
load(packages)
```

# Introducton

# Single Compartment SIR Model


We have $t = \frac{\tau}{\gamma}$, $s = S/N$, $i = I/N$,
\begin{equation}
\dot{s}= -R_\tau si,
\end{equation} where here $\dot{X}=dX/d\tau$, $R_0= \beta_\max/\gamma$,
$R_\tau=\beta(\tau)/\gamma$ and $R_{t}=R_\tau s$.


```{=tex}
\begin{equation}
\dot{i}=i (R_\tau s-1) = i (R_t-1)
\label{eq:didt}
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_s}= R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_i}= R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}
```

The full Lagrangian is:

\begin{equation}
\mathcal{L} = -c \ln\left(\frac{R_{\tau}}{R_0}\right)+c\left(\frac{R_{\tau}}{R_0}\right)+R_{\tau} si +
\lambda_s\left(\frac{d}{d \tau}{s} +R_{\tau} si\right)+
\lambda_i\left(\frac{d}{d \tau}i-R_{\tau}si+i \right)
\end{equation}

Apply equation 1 $x_i = s$

\begin{equation}
\dot{\lambda}_s = R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}

Apply equation 1 $x_i = i$:
\begin{equation}
\dot{\lambda}_i = R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}

Apply equation 1 $x_i = R_{\tau}$:

\begin{equation}
R_{\tau} = \frac{R_0c}{c+R_0si(1+\lambda_s-\lambda_i)}
\end{equation}

## Implementation

```{r}
oneComparmentSIRmyopic <- function(x, y, parms) {

    with(as.list(c(y, parms)),{
        s <- y[1] 
        i <- y[2] 
       
        
        R_tau = R_0*c/(c+R_0*s*i)
        
        ds <- -R_tau *s *i
        
        di <- i * (R_tau*s-1)

    return(list(c(ds,di)))
    })
}
```




Functions for solving boundary value problems include bvptwp, bvpcol,
and bvpshoot

```{r}
#i0 <- 0.001
#yini = c(1-i0,i0,NA,NA)
#yend = c(NA,NA,0,0)
#t <- seq(0,8,1/7) ## 112 days in steps of 2 days assuming 1/gamma=14 days

#t.vec <- seq(8,64,1) # from 112 days  to just under 2.5 years of social distancing
#c.vec <- seq(0.1,4.1,0.5)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  seq(1.5,5,0.5)
#i0.vec <- seq(1,51,5)*1e-4


#t.vec <- 4^seq(2,5,1)/2 # from 112 days  to just under 2.5 years of social distancing
#c.vec <- .25^seq(2,5,1)*2 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(1,51,10)*1e-4

#t.vec <- seq(8,64,28) # from 112 days  to just under 2.5 years of social distancing
#c.vec <-  seq(0.1,12.1,4)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(51,1,-10)*1e-3
```

```{r}
sol.explore <- read.csv("INTERMEDIATE/sol.explore-step11.csv")
exper.design <- read.csv("INTERMEDIATE/exper.design-step11.csv")
```





```{r}
param.vec <- c("tFinal","c","R0","i0")
exper.design.std <- exper.design
params.scaled <- scale(exper.design[,param.vec])
exper.design.std[,param.vec] <- params.scaled
```

```{r}
dist.matrix <- dist(params.scaled,upper=TRUE)
```

```{r}
dist.matrix <- as.matrix(dist.matrix)
```

```{r}
exper.design.s <- exper.design[exper.design$twp.success==TRUE,]
exper.design.u <- exper.design[exper.design$c %in% c(.025, 0.05, 0.1, 0.2, 0.4),]
exper.design.std.s <- exper.design.std[exper.design$twp.success==TRUE,]
exper.design.std.u <- exper.design.std[exper.design$twp.success==FALSE,]
```

```{r}
exper.design$twp.success11 <- NA
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
print(Sys.time())
for (case.run in exper.design.u$case.id){


  
   print(paste0("Case run is",case.run)) 
   case.id <- exper.design$case.id[case.run]
   tFinal <- exper.design$tFinal[case.run]
   c <- exper.design$c[case.run]
   R_0 <- exper.design$R0[case.run]
   i0 <- exper.design$i0[case.run]
    t <- seq(0,tFinal,1/7)
    t.temp <- t
    yini = c(1-i0,i0)
    solved <- FALSE
    
    print(paste("tFinal of case is",tFinal))
        print(paste("R0 is",R_0))
        print(paste("c is",c))
print(paste("c*tFinal is",tFinal*c))
  same.c <- exper.design.s[exper.design.s$tFinal <= tFinal,]
 # dists <- dist.matrix[same.c$case.id,case.run]
#  cases <- names(sort(dists)[1:5])
#  same.c <- subset(exper.design.s,case.id %in% cases)




      t <- seq(0,tFinal,1/7)
sol <- ode(y = yini,times = t,func = oneComparmentSIRmyopic,parms = list(R_0 = R_0,c=c))
    


solved <- TRUE    
    if(solved==TRUE){
      sol.explore <- subset(sol.explore,case.id!=case.run)
    sol <- as.data.frame(sol)
    sol <- sol[,1:3]
    names(sol) <- c("x",as.character(1:2))
    
    sol <- as.data.frame(sol) %>% 
    rename(tau=x,s='1',i='2') %>% 
    mutate(R_tau = R_0*c/(c+R_0*s*i),
           Rt= R_tau*s,
           tFinal=tFinal,
           case.id = case.id)
    
    sol.explore  <- bind_rows(sol.explore,sol)
    }
     }
print(Sys.time())
write.csv(sol.explore,"INTERMEDIATE/sol.explore-myopic.csv",row.names = FALSE)

```

