---
title: "COVID-19 Mitigation vs. Suppression Social Distancing"
author:
- name: Sarah Nowak, Pedro de Lima, Raffaele Vardavas 
  affiliation: RAND Corporation
#bibliography: ../../WriteUp/COVID-19.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: html_notebook
output:
  rmdformats::readthedown
pandoc_args: --natbib
biblio-style: plain
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
# Needed packages to perform the analysis
load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
} 

packages <- c("knitr","kableExtra","rmdformats","rmarkdown","dplyr","here","DiagrammeR",
              "ggplot2","scales","ggthemes","bvpSolve","deSolve","class","pracma")
load(packages)
```

# Introducton

# Single Compartment SIR Model


We have $t = \frac{\tau}{\gamma}$, $s = S/N$, $i = I/N$,
\begin{equation}
\dot{s}= -R_\tau si,
\end{equation} where here $\dot{X}=dX/d\tau$, $R_0= \beta_\max/\gamma$,
$R_\tau=\beta(\tau)/\gamma$ and $R_{t}=R_\tau s$.


```{=tex}
\begin{equation}
\dot{i}=i (R_\tau s-1) = i (R_t-1)
\label{eq:didt}
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_s}= R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_i}= R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}
```

The full Lagrangian is:

\begin{equation}
\mathcal{L} = -c \ln\left(\frac{R_{\tau}}{R_0}\right)+c\left(\frac{R_{\tau}}{R_0}\right)+R_{\tau} si +
\lambda_s\left(\frac{d}{d \tau}{s} +R_{\tau} si\right)+
\lambda_i\left(\frac{d}{d \tau}i-R_{\tau}si+i \right)
\end{equation}

Apply equation 1 $x_i = s$

\begin{equation}
\dot{\lambda}_s = R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}

Apply equation 1 $x_i = i$:
\begin{equation}
\dot{\lambda}_i = R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}

Apply equation 1 $x_i = R_{\tau}$:

\begin{equation}
R_{\tau} = \frac{R_0c}{c+R_0si(1+\lambda_s-\lambda_i)}
\end{equation}

## Implementation

```{r}
oneComparmentSIRmyopic <- function(x, y, parms) {
    with(as.list(c(y, parms)),{
        s <- y[1] 
        i <- y[2]
        cost <- y[3]
          if (x > delay) {
      lagged_values <- lagvalue(x - delay)
      s_d <- lagged_values[1]
      i_d <- lagged_values[2]
    } else {
      s_d <- s
      i_d <- i
    }
        
        R_tau = R_0*c/(c+R_0*s_d*i_d)
        
        ds <- -R_tau *s *i
        
        di <- i * (R_tau*s-1)
        
       dcost <- -c*log(R_tau/R_0)+c*(R_tau/R_0)+R_tau*s*i-c

    return(list(c(ds,di,dcost),R_tau=R_tau))
    })
}
```




Functions for solving boundary value problems include bvptwp, bvpcol,
and bvpshoot

```{r}
#i0 <- 0.001
#yini = c(1-i0,i0,NA,NA)
#yend = c(NA,NA,0,0)
#t <- seq(0,8,1/7) ## 112 days in steps of 2 days assuming 1/gamma=14 days

#t.vec <- seq(8,64,1) # from 112 days  to just under 2.5 years of social distancing
#c.vec <- seq(0.1,4.1,0.5)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  seq(1.5,5,0.5)
#i0.vec <- seq(1,51,5)*1e-4


#t.vec <- 4^seq(2,5,1)/2 # from 112 days  to just under 2.5 years of social distancing
#c.vec <- .25^seq(2,5,1)*2 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(1,51,10)*1e-4

#t.vec <- seq(8,64,28) # from 112 days  to just under 2.5 years of social distancing
#c.vec <-  seq(0.1,12.1,4)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(51,1,-10)*1e-3
```

```{r}
sol.explore <- read.csv("INTERMEDIATE/sol.explore-step11.csv")
exper.design <- read.csv("INTERMEDIATE/exper.design-round13.csv")
exper.design <- read.csv("INTERMEDIATE/compare.runs.csv")
exper.design <- exper.design[,c("case.id","R0","c","i0","tFinal")]
exper.design <- exper.design[!duplicated(exper.design),]
```





```{r}
param.vec <- c("tFinal","c","R0","i0")
exper.design.std <- exper.design
params.scaled <- scale(exper.design[,param.vec])
exper.design.std[,param.vec] <- params.scaled
```

```{r}
dist.matrix <- dist(params.scaled,upper=TRUE)
```

```{r}
dist.matrix <- as.matrix(dist.matrix)
```




```{r message=FALSE, warning=FALSE, paged.print=FALSE}
print(Sys.time())
exper.design <- exper.design[exper.design$R0!=6,]
exper.design0 <- exper.design
exper.design0$delay <- 0
exper.design1 <- exper.design
exper.design1$delay <- 1
exper.design2 <- exper.design
exper.design2$delay <- 2
exper.design3 <- exper.design
exper.design3$delay <- 3
exper.design4 <- exper.design
exper.design4$delay <- 4
exper.design <- rbind(exper.design0,exper.design1,exper.design2,exper.design3,exper.design4)
for (case.run in 1:length(exper.design$case.id)){


  
   print(paste0("Case run is",case.run)) 
   case.id <- exper.design$case.id[case.run]
   tFinal <- exper.design$tFinal[case.run]
   c <- exper.design$c[case.run]
   R_0 <- exper.design$R0[case.run]
   i0 <- exper.design$i0[case.run]
   delay <- exper.design$delay[case.run]
    t <- seq(0,tFinal,1/7)
    t.temp <- t
    yini = c(1-i0,i0,0)
    solved <- FALSE
    
    print(paste("tFinal of case is",tFinal))
        print(paste("R0 is",R_0))
        print(paste("c is",c))
print(paste("delay is",delay))



      t <- seq(0,tFinal,1/7)
sol <- dede(y = yini,times = t,func = oneComparmentSIRmyopic,parms = list(R_0 = R_0,c=c,delay=delay))

solved <- TRUE    
    
    sol <- as.data.frame(sol)
    sol <- sol[,1:5]
    names(sol) <- c("tau","s","i","cost","R_tau")
    
    sol <- as.data.frame(sol) %>% 
    mutate(Rt= R_tau*s,
           tFinal=tFinal,
           case.id = case.id,
           R0=R_0,
           c=c,
           delay=delay,
           i0=i0)
    if(case.run==1){
      sol.explore <- sol
      } else {
    sol.explore  <- bind_rows(sol.explore,sol)
      }
}

print(Sys.time())
write.csv(sol.explore,paste0("INTERMEDIATE/sol.explore-myopic-delay.csv"),row.names = FALSE)
write.csv(exper.design,"INTERMEDIATE/exper.design-delayround13.csv",row.names = FALSE)

```

