---
title: "COVID-19 Mitigation vs. Suppression Social Distancing"
author:
- name: Sarah Nowak, Pedro de Lima, Raffaele Vardavas 
  affiliation: RAND Corporation
#bibliography: ../../WriteUp/COVID-19.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: html_notebook
output:
  rmdformats::readthedown
pandoc_args: --natbib
biblio-style: plain
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
# Needed packages to perform the analysis
load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
} 

packages <- c("knitr","kableExtra","rmdformats","rmarkdown","dplyr","here","DiagrammeR",
              "ggplot2","scales","ggthemes","bvpSolve","deSolve")
load(packages)
```

# Introducton

# Single Compartment SIR Model


We have $t = \frac{\tau}{\gamma}$, $s = S/N$, $i = I/N$,
\begin{equation}
\dot{s}= -R_\tau si,
\end{equation} where here $\dot{X}=dX/d\tau$, $R_0= \beta_\max/\gamma$,
$R_\tau=\beta(\tau)/\gamma$ and $R_{t}=R_\tau s$.


```{=tex}
\begin{equation}
\dot{i}=i (R_\tau s-1) = i (R_t-1)
\label{eq:didt}
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_s}= R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_i}= R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}
```

The full Lagrangian is:

\begin{equation}
\mathcal{L} = -c \ln\left(\frac{R_{\tau}}{R_0}\right)+c\left(\frac{R_{\tau}}{R_0}\right)+R_{\tau} si +
\lambda_s\left(\frac{d}{d \tau}{s} +R_{\tau} si\right)+
\lambda_i\left(\frac{d}{d \tau}i-R_{\tau}si+i \right)
\end{equation}

Apply equation 1 $x_i = s$

\begin{equation}
\dot{\lambda}_s = R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}

Apply equation 1 $x_i = i$:
\begin{equation}
\dot{\lambda}_i = R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}

Apply equation 1 $x_i = R_{\tau}$:

\begin{equation}
R_{\tau} = \frac{R_0c}{c+R_0si(1+\lambda_s-\lambda_i)}
\end{equation}

## Implementation

```{r}
oneComparmentSIR <- function(x, y, parms) {

    with(as.list(c(y, parms)),{
        s <- y[1] 
        i <- y[2] 
        lambda_s <- y[3]
        lambda_i <- y[4]
        
        R_tau = R_0*c/(c+R_0*s*i*(1+lambda_s-lambda_i))
        
        ds <- -R_tau *s *i
        
        di <- i * (R_tau*s-1)
        
        dlambda_s <- R_tau*i*(1+lambda_s-lambda_i)
        
        dlambda_i <- R_tau*s*(1+lambda_s-lambda_i)+lambda_i

    return(list(c(ds,di,dlambda_s,dlambda_i)))
    })
}
```


Functions for solving boundary value problems include bvptwp, bvpcol,
and bvpshoot

```{r}
#i0 <- 0.001
#yini = c(1-i0,i0,NA,NA)
#yend = c(NA,NA,0,0)
#t <- seq(0,8,1/7) ## 112 days in steps of 2 days assuming 1/gamma=14 days

#t.vec <- seq(8,64,1) # from 112 days  to just under 2.5 years of social distancing
#c.vec <- seq(0.1,4.1,0.5)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  seq(1.5,5,0.5)
#i0.vec <- seq(1,51,5)*1e-4


#t.vec <- 4^seq(2,5,1)/2 # from 112 days  to just under 2.5 years of social distancing
#c.vec <- .25^seq(2,5,1)*2 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(1,51,10)*1e-4

#t.vec <- seq(8,64,28) # from 112 days  to just under 2.5 years of social distancing
#c.vec <-  seq(0.1,12.1,4)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(51,1,-10)*1e-3
```

```{r}
sol.explore <- read.csv("INTERMEDIATE/sol.explore-step6.csv")
exper.design <- read.csv("INTERMEDIATE/exper.design-round2.csv")
```

```{r}
exper.design.s <- exper.design[exper.design$twp.success==TRUE,]
exper.design.u <- exper.design[exper.design$twp.success==FALSE,]
```

```{r}

exper.design <- exper.design[order(exper.design$case.id),]
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
print(Sys.time())
for (case.run in exper.design.u$case.id){
  print(case.run)

  
   case.id <- exper.design$case.id[case.run]
   print(paste("case run is case id?",case.id == case.run))
   tFinal <- exper.design$tFinal[case.run]
   c <- exper.design$c[case.run]
   R_0 <- exper.design$R0[case.run]
   i0 <- exper.design$i0[case.run]
   
     same.tFinal <- subset(exper.design.s,tFinal==exper.design.u[which(exper.design.u$case.id == case.run),"tFinal"]&i0==exper.design.u[which(exper.design.u$case.id == case.run),"i0"]) 
  
  if(dim(same.tFinal)[1]==0){
    same.tFinal <- exper.design.s[sample(1:dim(exper.design.s)[1],20),]
  }
   print(tFinal)
    t <- seq(0,tFinal,1/7)
    yini = c(1-i0,i0,NA,NA)
    yend = c(NA,NA,0,0)
    solved <- FALSE
    for(case.id.c in same.tFinal$case.id){
      print(case.id.c)
     if(solved != TRUE){
      guess <- subset(sol.explore,case.id == case.id.c)
      if(length(t) < dim(guess)[1]){
        guess <- guess[1:length(t),]
      } else if(length(t) > dim(guess)[1]){
        guess[(dim(guess)[1]+1):length(t),]<- guess[dim(guess)[1],]
      }
    # Here we try the two-point method. Note we silenced the error so if it doesn't converge
    # we use the shooting method solution. 
    res <- try(sol <- bvptwp(yini = yini,yend=yend,x=t,parms = list(R_0 = R_0,c=c),
                         func = oneComparmentSIR,xguess = t,yguess = t(guess[,c("s","i","lambda_s","lambda_i")]),nmax=4000),
               silent=TRUE)
    exper.design[which(exper.design$case.id==case.run),"twp.success"] <- !inherits(res,"try-error")

    if(!inherits(res, "try-error")){
   solved <- TRUE
    }
    
    if(solved==TRUE){
      sol.explore <- subset(sol.explore,case.id!=case.run)
    sol <- as.data.frame(sol)
    sol <- sol[,1:5]
    names(sol) <- c("x",as.character(1:4))
    
    sol <- as.data.frame(sol) %>% 
    rename(tau=x,s='1',i='2',lambda_s='3',lambda_i='4') %>% 
    mutate(R_tau = R_0*c/(c+R_0*s*i*(1+lambda_s-lambda_i)),
           Rt= R_tau*s,
           tFinal=tFinal,
           case.id = case.id)
    
    sol.explore  <- bind_rows(sol.explore,sol)
    print(paste0("case id in case id c?",case.id %in% sol.explore$case.id))
    }
     }
    }
    print(solved)}
print(Sys.time())
write.csv(sol.explore,"INTERMEDIATE/sol.explore-step7.csv")
write.csv(exper.design,"INTERMEDIATE/exper.design-step7.csv")

```

