---
title: "COVID-19 Mitigation vs. Suppression Social Distancing"
author:
- name: Sarah Nowak, Pedro de Lima, Raffaele Vardavas 
  affiliation: RAND Corporation
#bibliography: ../../WriteUp/COVID-19.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: html_notebook
output:
  rmdformats::readthedown
pandoc_args: --natbib
biblio-style: plain
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
remove(list = ls())
# Needed packages to perform the analysis
load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
} 

packages <- c("knitr","kableExtra","rmdformats","rmarkdown","dplyr","here","DiagrammeR",
              "ggplot2","scales","ggthemes","bvpSolve","deSolve")
load(packages)
```


# Introducton



# Single Compartment SIR Model

Euler-Lagrange Equation:
\begin{equation}
\frac{d}{dt} \frac{\partial \cal{L}}{\partial {\dot{x_i}}}-\frac{\partial \cal{L}}{\partial x_i}
\end{equation}
 which becomes:
\begin{equation}
\frac{d}{dt} \frac{\partial \cal{L}}{\partial {\dot{x_i}}}=\frac{\partial \cal{L}}{\partial x_i}
\label{eq:EL}
\end{equation}


\begin{equation}
\dot{S}= b-\frac{\beta IS}{N}
\end{equation}



\begin{equation}
\dot{I}=\frac{\beta IS}{N}-\gamma I
\label{eq:didt}
\end{equation}

\begin{equation}
\dot{R}=\gamma I
\end{equation}

Cost function:
\begin{equation}
-a\rm{ln}({\beta})+c\beta+\frac{\beta IS}{N}
\label{eq:cost}
\end{equation}

The full Lagrangian, including constraints for the equations of motion is:
\begin{equation}
\begin{split}
\mathcal{L} = -a \rm {ln}({\beta})+c\beta+\frac{\beta IS}{N} +\\
\lambda_S(\dot{S} +\frac{\beta IS}{N})+\\
\lambda_I(\dot{I}-\frac{\beta IS}{N}+\gamma I)
\end{split}
\end{equation}
We leave out the constraint for $R$ because it doesn't factor into the cost equation.

Apply equation 1 $x_i = S$

\begin{equation}
\frac{d}{dt}\left(\lambda_s \right)=\left(\frac{\beta I}{N}+\lambda_S \frac{\beta I}{N} - \lambda_I \frac{\beta I}{N}\right)
\end{equation}

\begin{equation}
\dot{\lambda}_S = \frac{\beta I}{N}(1 + \lambda_S-\lambda_I)
\end{equation}

Apply equation 1 $x_i = I$:
\begin{equation}
\dot{\lambda}_I = \frac{\beta S}{N}(1+\lambda_S-\lambda_I)+\gamma \lambda_I
\end{equation}

Apply equation 1 $x_i = \beta$:
\begin{equation}
-\frac{a}{\beta}+c+\frac{IS}{N}+\lambda_S \frac{IS}{N}-\lambda_I \frac{IS}{N}
\end{equation}
which becomes
\begin{equation}
\frac{a}{\beta} = c + \frac{IS}{N}(1+\lambda_S-\lambda_I)
\end{equation}

\begin{equation}
\beta = \frac{a}{c+\frac{IS}{N}(1+\lambda_S-\lambda_I)}
\end{equation}


When the infection rate is very low, we expect $\beta$ to be:
\begin{equation}
\frac{a}{c}
\end{equation}

RAFF: {Seems to me that the problem is only relevant if $R_0>1$. If $R_0\sim 1$ an $R_0>1$, then $IS/N\sim I$ which is small. Hence since $R_0=\beta/\gamma$ we have $a>c\gamma$}


## Implementation 

```{r}
oneComparmentSIR <- function(x, y, parms) {

    with(as.list(c(y, parms)),{
        S <- y[1] 
        I <- y[2] 
        LambdaS <- y[3]
        LambdaI <- y[4]
        
        beta = a/(c+(I*S)/N*(1+LambdaS-LambdaI))
        
        dS <- -beta*S*I/N
        
        dI <- beta*S*I/N-gamma*I
        
        dLambdaS <- beta*I/N*(1+LambdaS-LambdaI)
        
        dLambdaI <- beta*S/N*(1+LambdaS-LambdaI)+gamma*LambdaI


     

    return(list(c(dS,dI,dLambdaS,dLambdaI)))
    })
}
```

```{r}
N <- 1000
R0 <- 2.5
gamma <- 1/14
beta0 <- R0*gamma
c <- 2e1
a <- beta0*c
```

Functions for solving boundary value problems include bvptwp, bvpcol, and bvpshoot

```{r}
I0 <- 30
yini = c(N-I0,I0,NA,NA)
yend = c(NA,NA,0,0)
t <- seq(0,50,2)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
sol.explore <- NULL
for (tFinal in c(50,100,200,400,800)){
    t <- seq(0,tFinal,tFinal/25)
    
    capture.output(sol <- bvpshoot(yini = yini,yend=yend,x=t,parms = list(a = a,c=c,N=N),func = oneComparmentSIR))

    sol <- as.data.frame(sol) %>% 
    rename(day=x,S='1',I='2',LambdaS='3',LambdaI='4') %>% 
    mutate(R0=a/(c+(I*S)/N*(1-LambdaS-LambdaI))/gamma,
           Rt= R0*(S/N),
           tFinal=tFinal)
    
    sol.explore  <- bind_rows(sol.explore,sol)
    
}
```

```{r}

testing <- sol.explore %>% 
    select(day, S, tFinal) %>% 
  tidyr::gather("agg_variable", "agg_value", -day, -tFinal)

pS<-ggplot(testing) +
  geom_line(aes(
    x = day,
    y = agg_value,
    group = tFinal,
    colour = tFinal
  ),
  size = 1.5) +
  facet_wrap(~ agg_variable)+
  xlab("day") +
  ylab("N") +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = trans_format("log10", math_format(10 ^ .x))
  # ) +
  guides(linetype = T) +
  theme_economist() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16) ,
    title = element_text(size = 18) ,
    strip.text = element_text(size = 12) ,
    axis.title.x = element_text(size = 18) ,
    axis.title.y = element_text(size = 18),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )

testing <- sol.explore %>% 
    select(day, I, tFinal) %>% 
  tidyr::gather("agg_variable", "agg_value", -day, -tFinal)

pI<-ggplot(testing) +
  geom_line(aes(
    x = day,
    y = agg_value,
    group = tFinal,
    colour = tFinal
  ),
  size = 1.5) +
  facet_wrap(~ agg_variable)+
  xlab("day") +
  ylab("N") +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = trans_format("log10", math_format(10 ^ .x))
  # ) +
  guides(linetype = T) +
  theme_economist() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16) ,
    title = element_text(size = 18) ,
    strip.text = element_text(size = 12) ,
    axis.title.x = element_text(size = 18) ,
    axis.title.y = element_text(size = 18),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )

testing <- sol.explore %>% 
    select(day, LambdaS, LambdaI, tFinal) %>% 
  tidyr::gather("agg_variable", "agg_value", -day, -tFinal)

pLambda<-ggplot(testing) +
  geom_line(aes(
    x = day,
    y = agg_value,
    group = tFinal,
    colour = tFinal
  ),
  size = 1.5) +
  facet_wrap(~ agg_variable)+
  xlab("day") +
  ylab("Lambda") +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = trans_format("log10", math_format(10 ^ .x))
  # ) +
  guides(linetype = T) +
  theme_economist() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16) ,
    title = element_text(size = 18) ,
    strip.text = element_text(size = 12) ,
    axis.title.x = element_text(size = 18) ,
    axis.title.y = element_text(size = 18),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )

testing <- sol.explore %>% 
    select(day, R0, Rt, tFinal) %>% 
  tidyr::gather("agg_variable", "agg_value", -day, -tFinal)

pR<-ggplot(testing) +
  geom_line(aes(
    x = day,
    y = agg_value,
    group = tFinal,
    colour = tFinal
  ),
  size = 1.5) +
  facet_wrap(~ agg_variable)+
  xlab("day") +
  ylab("Reproductive number") +
  # scale_y_log10(
  #   breaks = trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = trans_format("log10", math_format(10 ^ .x))
  # ) +
  guides(linetype = T) +
  theme_economist() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16) ,
    title = element_text(size = 18) ,
    strip.text = element_text(size = 12) ,
    axis.title.x = element_text(size = 18) ,
    axis.title.y = element_text(size = 18),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )


print(pS)
print(pI)
print(pLambda)
print(pR)
```

# Individual, myopic optimum

\begin{equation}
\frac{d}{d \beta}\left[-a\rm{ln}({\beta})+c\beta+\beta\frac{ IS}{N}\right] = 0
\end{equation}
which becomes 
\begin{equation}
-\frac{a}{\beta}+c+\frac{IS}{N} = 0
\end{equation}

\begin{equation}
\frac{a}{\beta} = c+\frac{IS}{N}
\end{equation}

\begin{equation}
\beta = \frac{a}{c+\frac{IS}{N}}
\end{equation}

# Model with exponential discounting

Here, we add exponential discounting to the single-compartment SIR model. 
Adding exponential discounting to the cost function in Eq. \ref{eq:cost}, we have:
\begin{equation}
e^{-\alpha t}\left(-a\rm{ln}({\beta})+c\beta+\frac{\beta IS}{N}\right)
\end{equation}
\note[rv]{I see why you adde the ddscounting on the cost function. But for large $\alpha$ we do not recover the Myopic formulation. Might the discounting be placed on the SIR constraints?}
 The full Lagrangian equation becomes:
\begin{equation}
\begin{split}
\mathcal{L} =e^{-\alpha t} \left(-a \rm {ln}({\beta})+c\beta+\frac{\beta IS}{N}\right) +\\
\lambda_S(\dot{S} +\frac{\beta IS}{N})+\\
\lambda_I(\dot{I}-\frac{\beta IS}{N}+\gamma I)
\end{split}
\end{equation}

Applying Eq. \ref{eq:EL} with $x_i = S$, we have 
\begin{equation}
\dot{\lambda}_S = e^{-\alpha t}(\frac{\beta I}{N})+\lambda_S \frac{\beta I}{N}-\lambda_I \frac{\beta I}{N}
\end{equation}
which becomes
\begin{equation}
\dot{\lambda}_S = \frac{\beta I}{N}\left( e^{-\alpha t}+\lambda_S-\lambda_I \right)
\end{equation}
then, applying Eq. \ref{eq:EL} with $x_i = I$, we get:
\begin{equation}
\dot{\lambda}_I = e^{-\alpha t}\frac{\beta S}{N}+\lambda_S \frac{\beta S}{N}+\lambda_I(\gamma-\frac{\beta S}{N})
\end{equation}
which becomes
\begin{equation}
\dot{\lambda}_I = \frac{\beta S}{N}(e^{-\alpha t}+\lambda_S-\lambda_I)+\gamma \lambda_I
\end{equation}
then, applying Eq. \ref{eq:EL} with $x_i = \beta$, we get:
\begin{equation}
0=e^{-\alpha t}(-\frac{a}{\beta}+c+\frac{IS}{N})+(\lambda_S-\lambda_I)\frac{IS}{N}
\end{equation}
\begin{equation}
\frac{a}{\beta} = c+\frac{IS}{N}+e^{\alpha t}(\lambda_S-\lambda_I)\frac{IS}{N}
\end{equation}

\begin{equation}
\frac{a}{\beta} = c+\frac{IS}{N}(1+e^{\alpha t}(\lambda_S-\lambda_I))
\end{equation}

\begin{equation}
\beta = \frac{a}{c+\frac{IS}{N}(1+e^{\alpha t}\lambda_S-e^{\alpha t}\lambda_I)}
\end{equation}

# Multi-Compartment SIR Model

\begin{equation}
\dot{S_i} = -\frac{1}{N}\sum_j{\beta_{ij}I_jS_i}
\end{equation}

\begin{equation}
\dot{I_i} = \frac{1}{N}\sum_j{\beta_{ij}I_jS_i}-\gamma I_i
\end{equation}

Cost function:
\begin{equation}
\mathcal{C} =\sum_i N_i\left( -a_{ij}\rm{ln}({\beta_{ij}})+c\beta_{ij}+\frac{\beta_{ij}{I_jS_i}}{N_j}\right)
\end{equation}
Lagrangian:
\begin{equation}
\mathcal{L} = \sum_i N_i\left( -a_{ij}\rm{ln}({\beta_{ij}})+c\beta_{ij}+\frac{\beta_{ij}{I_jS_i}}{N}\right)+
\sum_i \lambda_{S,i}\left( \dot{S}_i+\frac{1}{N}\sum_j{\beta_{ij}I_jS_i}\right)+\sum_i \lambda_{I,i}\left(\dot{I}_i -\frac{\beta_{ij}I_jS_i}{N}+\gamma I_i  \right)
\end{equation}



