---
title: "COVID-19 Mitigation vs. Suppression Social Distancing"
author:
- name: Sarah Nowak, Pedro de Lima, Raffaele Vardavas 
  affiliation: RAND Corporation
#bibliography: ../../WriteUp/COVID-19.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: html_notebook
output:
  rmdformats::readthedown
pandoc_args: --natbib
biblio-style: plain
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
# Needed packages to perform the analysis
load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
} 

packages <- c("knitr","kableExtra","rmdformats","rmarkdown","dplyr","here","DiagrammeR",
              "ggplot2","scales","ggthemes","bvpSolve","deSolve","plyr")
load(packages)
```

# Introducton

# Single Compartment SIR Model


We have $t = \frac{\tau}{\gamma}$, $s = S/N$, $i = I/N$,
\begin{equation}
\dot{s}= -R_\tau si,
\end{equation} where here $\dot{X}=dX/d\tau$, $R_0= \beta_\max/\gamma$,
$R_\tau=\beta(\tau)/\gamma$ and $R_{t}=R_\tau s$.


```{=tex}
\begin{equation}
\dot{i}=i (R_\tau s-1) = i (R_t-1)
\label{eq:didt}
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_s}= R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}
```
```{=tex}
\begin{equation}
\dot{\lambda_i}= R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}
```

The full Lagrangian is:

\begin{equation}
\mathcal{L} = -c \ln\left(\frac{R_{\tau}}{R_0}\right)+c\left(\frac{R_{\tau}}{R_0}\right)+R_{\tau} si +
\lambda_s\left(\frac{d}{d \tau}{s} +R_{\tau} si\right)+
\lambda_i\left(\frac{d}{d \tau}i-R_{\tau}si+i \right)
\end{equation}

Apply equation 1 $x_i = s$

\begin{equation}
\dot{\lambda}_s = R_{\tau} i (1 + \lambda_s-\lambda_i)
\end{equation}

Apply equation 1 $x_i = i$:
\begin{equation}
\dot{\lambda}_i = R_{\tau}s(1+\lambda_s-\lambda_i)+\lambda_i
\end{equation}

Apply equation 1 $x_i = R_{\tau}$:

\begin{equation}
R_{\tau} = \frac{R_0c}{c+R_0si(1+\lambda_s-\lambda_i)}
\end{equation}

## Implementation

```{r}
oneComparmentSIR <- function(x, y, parms) {

    with(as.list(c(y, parms)),{
        s <- y[1] 
        i <- y[2] 
        lambda_s <- y[3]
        lambda_i <- y[4]
        
        R_tau = R_0*c/(c+R_0*s*i*(1+lambda_s-lambda_i))
        
        ds <- -R_tau *s *i
        
        di <- i * (R_tau*s-1)
        
        dlambda_s <- R_tau*i*(1+lambda_s-lambda_i)
        
        dlambda_i <- R_tau*s*(1+lambda_s-lambda_i)+lambda_i

    return(list(c(ds,di,dlambda_s,dlambda_i)))
    })
}
```


Functions for solving boundary value problems include bvptwp, bvpcol,
and bvpshoot

```{r}
i0 <- 0.001
yini = c(1-i0,i0,NA,NA)
yend = c(NA,NA,0,0)
t <- seq(0,8,1/7) ## 112 days in steps of 2 days assuming 1/gamma=14 days

t.vec <- seq(8,64,1) # from 112 days  to just under 2.5 years of social distancing
c.vec <- seq(0.1,4.1,0.5)/64 # chosen such that  mean(c.vec%o%t.vec)~1
R0.vec <-  seq(1.5,5,0.5)
i0.vec <- seq(1,51,5)*1e-4


t.vec <- 2^seq(3,7,1) # from 112 days  to just under 2.5 years of social distancing
c.vec <- 0.025*2^seq(0,4,1) # chosen such that  mean(c.vec%o%t.vec)~1
R0.vec <-  c(1.5,2.5,5)
i0.vec <- seq(1,51,25)*1e-4

#t.vec <- seq(8,64,28) # from 112 days  to just under 2.5 years of social distancing
#c.vec <-  seq(0.1,12.1,4)/64 # chosen such that  mean(c.vec%o%t.vec)~1
#R0.vec <-  c(1.5,5)
#i0.vec <- seq(51,1,-10)*1e-3

exper.design <- base::expand.grid(tFinal = t.vec, c=c.vec, R0=R0.vec, i0= i0.vec) %>% 
  dplyr::mutate(case.id=row_number())

exper.design %>% 
  mutate(prod=tFinal*c) %>% 
  filter(prod<1) 

exper.design %>% 
  mutate(prod=tFinal*c) %>% 
  filter(prod>1) 
use.steps <- FALSE
```

```{r}
sol.explore6 <- read.csv("INTERMEDIATE/sol.explore-step6.csv")
exper.design6 <- read.csv("INTERMEDIATE/exper.design-step6.csv")
```

```{r}
exper.design$case.id <- exper.design$case.id + dim(exper.design6)[1]
exper.design$twp.success <- FALSE
```

```{r}
exper.design7 <- rbind.fill(exper.design,exper.design6) 
```

```{r}
write.csv(exper.design7,"INTERMEDIATE/exper.design-round2.csv")

```
